---
title: "GCA Network"
output: html_notebook
---

Time to load up my two favorite libraries again.

```{r}
library(tidyverse)
library(worldfootballR)
```


For the graph we will be making, we will also need igraph.

```{r}
library(igraph)
```


We will be using the 2017-18 Liverpool season to highlight the benefits of goal creation networks.

```{r}
lpool_url <- 'https://fbref.com/en/squads/822bd0ba/2017-2018/Liverpool-Stats'
goal_log <- fb_team_goal_logs(team_urls = lpool_url, for_or_against="for")
```


After looking through the goal_log, it's evident to me that the Champions' League playoff stage and the FA Cup are two tournaments where goal-creating actions are not tracked. Therefore we will filter out these tournaments.

```{r}
filtered_goal_log <- goal_log |>
  filter(!(goal_log$Round == 'Play-off round' | goal_log$Comp == 'FA Cup'))
```


The only columns needed will be Scorer, GCA1, and GC2.

```{r}
selected_goal_log <- filtered_goal_log |>
  select(Scorer, GCA1, GCA2)
```


We should also remove scorers that were own goals, as they are not Liverpool players.

```{r}
filtered_scorers_goal_log <- selected_goal_log |>
  mutate(Scorer = ifelse(grepl('(OG)', Scorer), "", Scorer))
```


Now we can build the graph.

```{r}
g <- make_empty_graph(directed = FALSE)
for (i in 1:nrow(filtered_scorers_goal_log)) {
  players <- filtered_scorers_goal_log[i, c("Scorer", "GCA1", "GCA2")]
  players <- players[players != ""]
  
  for (player in players) {
    if (!(player %in% V(g)$name)) {
      g <- add_vertices(g, 1, name = player)
    }
  }
  
  if (length(players) > 1) {
    edges <- unlist(combn(players, 2), use.names = FALSE)
    g <- add_edges(g, edges)
  }
}
```


With this graph we can see who the three most important goal creators are by three different methods: degree,...

```{r}
degree_df <- data.frame(
  vertex = V(g)$name,
  degree = degree(g)
)

degree_df[order(-degree_df$degree)[1:3], ]
```


...eigenvector,...

```{r}
eigenvector_df <- data.frame(
  vertex = V(g)$name,
  eigenvector = eigen_centrality(g)$vector
)

eigenvector_df[order(-eigenvector_df$eigenvector)[1:3], ]
```


...and closeness.

```{r}
closeness_df <- data.frame(
  vertex = V(g)$name,
  closeness = closeness(g)
)

closeness_df[order(-closeness_df$closeness)[1:3], ]
```


We can even create a visual of the graph that confirms what all three measures of centrality have told us: Mohamed Salah, Sadio Mané, and Roberto Firmino were the most integral to Liverpool's attack in the 2017-18 season.

```{r}
plot(g, vertex.label = V(g)$name, edge.color = "gray", vertex.size = 30)
```

