First step is to import the needed libraries.

```{r}
library(worldfootballR)
library(tidyverse)
library(mclust)
```

I then scrape all the data I need.

```{r}
big5_shooting <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'shooting', 
  team_or_player = 'player'
  )

big5_passing <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'passing', 
  team_or_player = 'player'
  )

big5_passing_types <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'passing_types', 
  team_or_player = 'player'
  )

big5_gca <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'gca', 
  team_or_player = 'player'
  )

big5_defense <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'defense', 
  team_or_player = 'player'
  )

big5_possession <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'possession', 
  team_or_player = 'player'
  )

big5_keepers <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'keepers', 
  team_or_player = 'player'
  )

big5_keepers_adv <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'keepers_adv', 
  team_or_player = 'player'
  )
```


Let's separate outfield players from goalies.
```{r}
big5_defense_gk <- big5_defense |>
  filter(Pos == "GK")
big5_defense_outfield <- big5_defense |>
  filter(Pos != "GK")

big5_gca_gk <- big5_gca |>
  filter(Pos == "GK")
big5_gca_outfield <- big5_gca |>
  filter(Pos != "GK")

big5_passing_gk <- big5_passing |>
  filter(Pos == "GK")
big5_passing_outfield <- big5_passing |>
  filter(Pos != "GK")

big5_passing_types_gk <- big5_passing_types |>
  filter(Pos == "GK")
big5_passing_types_outfield <- big5_passing_types |>
  filter(Pos != "GK")

big5_possession_gk <- big5_possession |>
  filter(Pos == "GK")
big5_possession_outfield <- big5_possession |>
  filter(Pos != "GK")

big5_shooting_gk <- big5_shooting |>
  filter(Pos == "GK")
big5_shooting_outfield <- big5_shooting |>
  filter(Pos != "GK")
```

Before we merge the different types of stats, we need to take care of the
players who played for multiple teams, as their data are separated. We'll also
remove all non-cumulative data in this step, as it not only simplifies this
step but it is also something we already were going to do.

```{r}
big5_defense_gk_no_dups <- big5_defense_gk |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Tkl_percent_Challenges,
    -Mins_Per_90
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_defense_outfield_no_dups <- big5_defense_outfield |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Tkl_percent_Challenges,
    -Mins_Per_90
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_gca_gk_no_dups <- big5_gca_gk |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -SCA90_SCA,
    -GCA90_GCA
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_gca_outfield_no_dups <- big5_gca_outfield |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -SCA90_SCA,
    -GCA90_GCA
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_keepers_no_dups <- big5_keepers |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -MP_Playing,
    -Starts_Playing,
    -Min_Playing,
    -Mins_Per_90,
    -Save_percent,
    -W,
    -D,
    -L,
    -CS_percent,
    -Save_percent_Penalty,
    -GA90
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_keepers_adv_no_dups <- big5_keepers_adv |>
  mutate(
    TotLen_Passes = AvgLen_Passes * (`Att (GK)_Passes` + Thr_Passes),
    TotLen_Goal = AvgLen_Goal * Att_Goal,
    TotDist_Sweeper = AvgDist_Sweeper * `#OPA_Sweeper`
  ) |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -`_per_90_Expected`,
    -Cmp_percent_Launched,
    -Launch_percent_Passes,
    -Launch_percent_Goal,
    -Stp_percent_Crosses,
    -PSxG_per_SoT_Expected,
    -`PSxG+_per__minus__Expected`,
    -`#OPA_per_90_Sweeper`,
    -AvgLen_Passes,
    -AvgLen_Goal,
    -AvgDist_Sweeper
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_passing_gk_no_dups <- big5_passing_gk |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -Cmp_percent_Total,
    -Cmp_percent_Short,
    -Cmp_percent_Medium,
    -Cmp_percent_Long
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_passing_outfield_no_dups <- big5_passing_outfield |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -Cmp_percent_Total,
    -Cmp_percent_Short,
    -Cmp_percent_Medium,
    -Cmp_percent_Long
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_passing_types_gk_no_dups <- big5_passing_types_gk |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_passing_types_outfield_no_dups <- big5_passing_types_outfield |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_possession_gk_no_dups <- big5_possession_gk |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -Succ_percent_Take,
    -Tkld_percent_Take
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_possession_outfield_no_dups <- big5_possession_outfield |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -Succ_percent_Take,
    -Tkld_percent_Take
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_shooting_gk_no_dups <- big5_shooting_gk |>
  mutate(
    TotDist_Standard = Dist_Standard * Sh_Standard
  ) |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -SoT_percent_Standard,
    -SoT_per_90_Standard,
    -Sh_per_90_Standard,
    -G_per_Sh_Standard,
    -G_per_SoT_Standard,
    -Dist_Standard,
    -npxG_per_Sh_Expected
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_shooting_outfield_no_dups <- big5_shooting_outfield |>
  mutate(
    TotDist_Standard = Dist_Standard * Sh_Standard
  ) |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -SoT_percent_Standard,
    -SoT_per_90_Standard,
    -Sh_per_90_Standard,
    -G_per_Sh_Standard,
    -G_per_SoT_Standard,
    -Dist_Standard,
    -npxG_per_Sh_Expected
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```


Now, we need to merge the data frames.

```{r}
big5_gk <- merge(
  big5_shooting_gk_no_dups, merge(
    big5_passing_gk_no_dups, 
    merge(
      big5_passing_types_gk_no_dups, 
      merge(
        big5_gca_gk_no_dups, 
        merge(
          big5_defense_gk_no_dups, 
          merge(
            big5_possession_gk_no_dups, 
            merge(
              big5_keepers_no_dups, 
              big5_keepers_adv_no_dups, 
              by = 'Url', 
              all = TRUE
            ), 
            by = 'Url', 
            all = TRUE
          ), 
          by = 'Url', 
          all = TRUE
        ), 
        by = 'Url', 
        all = TRUE
      ), 
      by = 'Url', 
      all = TRUE
    ), 
    by = 'Url', 
    all = TRUE
  ), 
  by = 'Url', 
  all = TRUE
)
```

```{r}
big5_outfield <- merge(
  big5_shooting_outfield_no_dups, merge(
    big5_passing_outfield_no_dups, 
    merge(
      big5_passing_types_outfield_no_dups, 
      merge(
        big5_gca_outfield_no_dups, 
        merge(
          big5_defense_outfield_no_dups, 
          big5_possession_outfield_no_dups,
          by = 'Url', 
          all = TRUE
          ),
        by = 'Url', 
        all = TRUE
        ), 
      by = 'Url', 
      all = TRUE
    ), 
    by = 'Url', 
    all = TRUE
  ), 
  by = 'Url', 
  all = TRUE
)
```


We'll replace all the NAs with 0s.

```{r}
big5_gk[is.na(big5_gk)] <- 0
big5_outfield[is.na(big5_outfield)] <- 0
```


Now that we have everything cleaned, we can normalize the data using
L1 normalization.

```{r}
l1_normalize <- function(x) {
  return(x / sum(abs(x)))
}
```

```{r}
row.names(big5_gk) <- big5_gk$Url
big5_gk$Url <- NULL

row.names(big5_outfield) <- big5_outfield$Url
big5_outfield$Url <- NULL
```

```{r}
big5_gk_normalized <- as.data.frame(t(apply(big5_gk, 1, l1_normalize)))
big5_outfield_normalized <- as.data.frame(t(apply(
  big5_outfield, 
  1, 
  l1_normalize
  )))
```


We will do another NA check, as it is possible for players to have only played
a minimal amount of time in a season, not charting any statistics. We do not
want that player in the data set, so we will remove them.
```{r}
big5_gk_normalized <- na.omit(big5_gk_normalized)
big5_outfield_normalized <- na.omit(big5_outfield_normalized)
```


We have a lot of dimensions. Too many. Let's reduce them using PCA.

For this to work, we need to remove any columns where no goalies recorded any
statistics at all.

```{r}
big5_gk_partial_pca <- big5_gk_normalized[, apply(
  big5_gk_normalized, 
  2, 
  var
  ) > 0]
```

```{r}
gk_pca_result <- prcomp(big5_gk_partial_pca, center = TRUE, scale. = TRUE)
big5_gk_pca <- predict(gk_pca_result, newdata = big5_gk_partial_pca)

outfield_pca_result <- prcomp(
  big5_outfield_normalized, 
  center = TRUE, 
  scale. = TRUE)
big5_outfield_pca <- predict(outfield_pca_result, 
                             newdata = big5_outfield_normalized)

summary(gk_pca_result)
summary(outfield_pca_result)
```


By viewing the summaries of the PCA results for the GK and Outfield data, we
can see that not all columns are of value when clustering. To help reduce
dimensionality, we should pick the least number of columns that gives us a
cumulative proportion of at least 0.95. For the GK data, that's up to PC43; for
the outfielders, that's up to PC49.
```{r}
big5_gk_pca_final <- big5_gk_pca[,1:43]
big5_outfield_pca_final <- big5_outfield_pca[,1:49]
```


Now, we can run the Gaussian Mixtures models.
```{r}
gk_gmm_model <- Mclust(big5_gk_pca_final)
big5_gk_clusters <- big5_gk |>
  mutate(Clusters = gk_gmm_model$classification)
```

```{r}
outfield_gmm_model <- Mclust(big5_outfield_pca_final)
big5_outfield_clusters <- big5_outfield |>
  rownames_to_column("temp") |>
  filter(temp %in% rownames(big5_outfield_normalized)) |>
  column_to_rownames("temp") |>
  mutate(Clusters = outfield_gmm_model$classification)
```


We got eight clusters from the outfield players and 3 from the goalkeepers. Good start, but I think we can get more out of each outfield cluster. Let's see if we can get more.
```{r}
cluster_models <- list()
cluster_results <- list()

unique_clusters <- unique(big5_outfield_clusters$Clusters)

for (cluster in unique_clusters) {
  # Subset the data for the current cluster
  cluster_data <- big5_outfield_clusters |>
    filter(Clusters == cluster) |>
    select(-Clusters)
  
  # Run GMM on the current cluster
  gmm_model <- Mclust(cluster_data)
  
  # Save the model to the list
  cluster_models[[as.character(cluster)]] <- gmm_model
  
  # Add the classification to the data
  cluster_result <- cluster_data |>
    mutate(Cluster = gmm_model$classification)
  
  # Save the results to the list
  cluster_results[[as.character(cluster)]] <- cluster_result
}
```


This generated 61 clusters in total, plus three more for goalkeepers! This is great, especially considering several (around 10) of the clusters are very small. I think having around 50 is good. Let's merge the outfielders back together, as they are currently separated by big cluster.
```{r}
big5_final <- bind_rows(cluster_results, .id = "ParentCluster")
big5_final
```