First step is to import the needed libraries.

```{r}
library(worldfootballR)
library(tidyverse)
```

I then scrape all the data I need.

```{r}
big5_shooting <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'shooting', 
  team_or_player = 'player'
  )

big5_passing <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'passing', 
  team_or_player = 'player'
  )

big5_passing_types <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'passing_types', 
  team_or_player = 'player'
  )

big5_gca <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'gca', 
  team_or_player = 'player'
  )

big5_defense <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'defense', 
  team_or_player = 'player'
  )

big5_possession <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'possession', 
  team_or_player = 'player'
  )

big5_keepers <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'keepers', 
  team_or_player = 'player'
  )

big5_keepers_adv <- fb_big5_advanced_season_stats(
  season_end_year = 2024, 
  stat_type = 'keepers_adv', 
  team_or_player = 'player'
  )
```

Let's create a player lookup. We will use Url as the IDs for this project,
but using names to find Urls would be much simpler.

```{r}
player_squad_url <- big5_passing |>
  select(Player, Squad, Url) |>
  distinct(Player, Squad, Url, .keep_all = TRUE)
```


Before we merge the data frames, we need to take care of the players who played
for multiple teams, as their data are separated. We'll also remove all
non-cumulative data in this step, as it not only simplifies this step but it is
also something we already were going to do.

```{r}
big5_defense_no_dups <- big5_defense |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Tkl_percent_Challenges,
    -Mins_Per_90
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_gca_no_dups <- big5_gca |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -SCA90_SCA,
    -GCA90_GCA
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_keepers_no_dups <- big5_keepers |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -MP_Playing,
    -Starts_Playing,
    -Min_Playing,
    -Mins_Per_90,
    -Save_percent,
    -W,
    -D,
    -L,
    -CS_percent,
    -Save_percent_Penalty,
    -GA90
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_keepers_adv_no_dups <- big5_keepers_adv |>
  mutate(
    TotLen_Passes = AvgLen_Passes * (`Att (GK)_Passes` + Thr_Passes),
    TotLen_Goal = AvgLen_Goal * Att_Goal,
    TotDist_Sweeper = AvgDist_Sweeper * `#OPA_Sweeper`
  ) |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -`_per_90_Expected`,
    -Cmp_percent_Launched,
    -Launch_percent_Passes,
    -Launch_percent_Goal,
    -Stp_percent_Crosses,
    -PSxG_per_SoT_Expected,
    -`PSxG+_per__minus__Expected`,
    -`#OPA_per_90_Sweeper`,
    -AvgLen_Passes,
    -AvgLen_Goal,
    -AvgDist_Sweeper
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_passing_no_dups <- big5_passing |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -Cmp_percent_Total,
    -Cmp_percent_Short,
    -Cmp_percent_Medium,
    -Cmp_percent_Long
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_passing_types_no_dups <- big5_passing_types |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_possession_no_dups <- big5_possession |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -Succ_percent_Take,
    -Tkld_percent_Take
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```

```{r}
big5_shooting_no_dups <- big5_shooting |>
  mutate(
    TotDist_Standard = Dist_Standard * Sh_Standard
  ) |>
  select(
    -Season_End_Year, 
    -Squad, 
    -Comp, 
    -Player,
    -Nation, 
    -Pos, 
    -Age, 
    -Born,
    -Mins_Per_90,
    -SoT_percent_Standard,
    -SoT_per_90_Standard,
    -Sh_per_90_Standard,
    -G_per_Sh_Standard,
    -G_per_SoT_Standard,
    -Dist_Standard,
    -npxG_per_Sh_Expected
    ) |>
  group_by(Url) |>
  summarize(across(everything(), sum))
```


Now, we need to merge the data frames.

```{r}
big5 <- merge(
  big5_shooting_no_dups, merge(
    big5_passing_no_dups, 
    merge(
      big5_passing_types_no_dups, 
      merge(
        big5_gca_no_dups, 
        merge(
          big5_defense_no_dups, 
          merge(
            big5_possession_no_dups, 
            merge(
              big5_keepers_no_dups, 
              big5_keepers_adv_no_dups, 
              by = 'Url', 
              all = TRUE
            ), 
            by = 'Url', 
            all = TRUE
          ), 
          by = 'Url', 
          all = TRUE
        ), 
        by = 'Url', 
        all = TRUE
      ), 
      by = 'Url', 
      all = TRUE
    ), 
    by = 'Url', 
    all = TRUE
  ), 
  by = 'Url', 
  all = TRUE
)
```


We'll replace all the NAs with 0s.

```{r}
big5[is.na(big5)] <- 0
```


Now that we have everything cleaned, we can normalize the data using
L1 normalization.

```{r}
l1_normalize <- function(x) {
  return(x / sum(abs(x)))
}
```

```{r}
row.names(big5) <- big5$Url
big5$Url <- NULL
```

```{r}
big5_normalized <- as.data.frame(t(apply(big5, 1, l1_normalize)))
```


Let's write a function that finds the 10 most similar players.

```{r}
similar_player_finder <- function(player, squad) {
  url <- player_squad_url$Url[player_squad_url$Player == player & 
                                player_squad_url$Squad == squad]
  distances <- as.matrix(
   dist(
     rbind(
       big5_normalized[rownames(big5_normalized) == url, ], big5_normalized
       )
     )
   )[1, -1]
  similar_players <- order(distances)[2:11]
  
  for (url in rownames(big5_normalized[similar_players, ])) {
   print(
     paste(
       player_squad_url$Player[player_squad_url$Url == url], 
       player_squad_url$Squad[player_squad_url$Url == url],
       sep = ', '
       )
     )
  }
}
```


Now let's try it! Let's find the 10 most similar players to Julián Álvarez.

```{r}
team <- 'Manchester City'
similar_player_finder('Julián Álvarez', team)
```

Now with Rodri. There are multiple players who go by Rodri, so the squad name is important.

```{r}
similar_player_finder('Rodri', team)
```

Lastly, with Kevin de Bruyne.

```{r}
similar_player_finder('Kevin De Bruyne', team)
```